---
title: "Mini Project 1"
---

```{r}
#| include: false

library(tidyverse)
library(tidytext)
library(textdata)
library(wordcloud)
library(wordcloud2)
library(viridis)
library(ggthemes)
library(gutenbergr)
```

```{r}
#| include: false
barack <- read_csv("https://proback.github.io/264_fall_2025/Data/tweets_potus.csv")
michelle <- read_csv("https://proback.github.io/264_fall_2025/Data/tweets_flotus.csv")

tweets <- bind_rows(barack |> 
                      mutate(person = "Barack"),
                    michelle |> 
                      mutate(person = "Michelle")) |>
  mutate(timestamp = ymd_hms(timestamp))

smart_stopwords <- get_stopwords(source = "smart")
words <- stringr::words
words
```

```{r}
#| include: false
tweets_words <- tweets |>
  mutate(
    date = ymd(str_extract(timestamp, "\\d{4}-\\d{2}-\\d{2}")),
    year = str_sub(date, 1, 4),
    month = str_sub(date, 6, 7)
  ) |>
  filter(date >= ymd("2015-5-18")) |>
  unnest_tokens(word, text) |>
  anti_join(smart_stopwords) |>
  filter(word %in% words) |>
  filter(!word %in% c("http", "https", "rt", "t.co"))

barack_words <- barack |>
  mutate(
    date = ymd(str_extract(timestamp, "\\d{4}-\\d{2}-\\d{2}")),
    year = str_sub(date, 1, 4),
    month = str_sub(date, 6, 7)
  ) |>
  unnest_tokens(word, text) |>
  anti_join(smart_stopwords) |>
  filter(word %in% words) |>
  filter(!word %in% c("http", "https", "rt", "t.co"))

michelle_words <- michelle |>
  mutate(
    date = ymd(str_extract(timestamp, "\\d{4}-\\d{2}-\\d{2}")),
    year = str_sub(date, 1, 4),
    month = str_sub(date, 6, 7)
  ) |>
  filter(date >= ymd("2015-5-18")) |>
  unnest_tokens(word, text) |>
  anti_join(smart_stopwords) |>
  filter(word %in% words) |>
  filter(!word %in% c("http", "https", "rt", "t.co"))
```

For one reason or another, the statistics gods have blessed us with data of all of President Barack Obama and First Lady Michelle Obama's entire tweet history during Obama's presidency. In this project we will explore the results of wrangling this data and utilizing it to create tons of interesting statistics, plots and graphs.

# Number of significant words tweeted per month

```{r}
#| echo: false
#| message: false
#| warning: false
barack_words |>
  count(year, month) |>
  ggplot() +
  geom_col(aes(x = month, y = n)) +
  facet_wrap(~year) +
  labs(
    title = "Number of significant words tweeted monthly by Barack Obama",
    subtitle = "(between May 2015 and November 2016)",
    caption = "**only contains words from the dplyr 'words' list**",
    x = "Month",
    y = "Number of words tweeted"
  )
```

```{r}
#| echo: false
#| message: false
#| warning: false
michelle_words |>
  count(year, month) |>
  ggplot() +
  geom_col(aes(x = month, y = n)) +
  facet_wrap(~year) +
  labs(
    title = "Number of significant words tweeted monthly by Michelle Obama",
    subtitle = "(between May 2015 and November 2016)",
    caption = "**only contains words from the dplyr 'words' list**",
    x = "Month",
    y = "Number of words tweeted"
  )
```

```{r, fig.alt = "The chart shown is a stacked bar graph showing the total number of significant words tweeted per month by both Barack and Michelle Obama during Obama's presidency. The x axis measures each month and spans from 05/2015 to 11/2016. The y axis measures the number of words tweeted and ranges from 10 to 155 words. The graph shows a significant variability in the number of words tweeted by month by both Barack and Michelle. It it also clear that Michelle was significantly more involved on Twitter during this time period compared to her husband since she continuously has the vast majority of significant words tweeted each month."}
#| echo: false
#| message: false
#| warning: false
tweets_words |>
  count(year, month, person) |>
  ggplot() +
  geom_col(aes(x = month, y = n, fill = person)) +
  facet_wrap(~year) +
  labs(
    title = "Combined number of significant words tweeted monthly\nbetween Barack and Michelle Obama",
    subtitle = "(between May 2015 and November 2016)",
    caption = "**only contains words from the dplyr 'words' list**",
    x = "Month",
    y = "Number of words tweeted",
    fill = "Person"
  )
```

Our story begins with simply counting how many significant words were tweeted by across each recorded month for a Barack-exclusive dataset, a Michelle-exclusive dataset, and a combined dataset. It's important to note that the term "significant words" can be defined as words that are not stopwords (ie: the, to, and, etc.), and words that are within dlpyr's word list package. This was done the because there were lots of stopwords and online links that were parsed into words during the data wrangling process that needed to be filtered out. Therefore, the words recorded in lots of the dataset are a very limited (but relevant) subset of the complete Twitter history of our former executive leaders.

We can see that there is a large variability within how much was tweeted by month for both Barack and Michelle, especially during 2015 which may call to the variety of everchanging issues and matter of which it is important to post about during each month. Both people also seemed to tweet much less during the following year at their own respective rate, which could potentially indicate a more relaxed, habitual usage of twitter as they get used to this new platform of media.

# Most common words by month

```{r}
#| echo: false
barack_words |>
  group_by(year, month) |>
  count(word) |>
  slice_max(order_by = n, with_ties = F)
```

```{r}
#| echo: false
#| message: false
#| warning: false
michelle_words |>
  group_by(year, month) |>
  count(word) |>
  slice_max(order_by = n, with_ties = F)
```

Next, we see the most common words by month recorded for both Barack and Michelle (since the majority of tweets during this time period come from Michelle, the combined dataset isn't used here). The words most commonly used by Barack emit a sense of political activism, initiative and inspiration--including great, change, health, people, history, etc.. On the other hand, almost of all of Michelle's most commonly used words by month were "lady" (possibly in self-reference to her title as first lady)--not at exciting at all.

# Most common word pairs

```{r}
#| echo: false
#| message: false
#| warning: false
barack |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  filter(bigram != "NA") |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word,
         !word1 %in% c("http", "https", "rt", "t.co"),
         !word2 %in% stop_words$word,
         !word2 %in% c("http", "https", "rt", "t.co")) |>
  count(word1, word2, sort = TRUE) |>
  print(n = 10)
```

```{r}
#| echo: false
#| message: false
#| warning: false
michelle |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  filter(bigram != "NA") |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word,
         !word1 %in% c("http", "https", "rt", "t.co"),
         !word2 %in% stop_words$word,
         !word2 %in% c("http", "https", "rt", "t.co")) |>
  count(word1, word2, sort = TRUE) |>
  print(n = 10)
```

After analyzing all of the most common individual words, I deemed it appropriate to take this data analysis to the next level and identify the most common pairs of words used by both Barack and Michelle in their tweets. Because there are so few examples to sample from, we will be counting from the total collection of pair words rather than sorting it by month. The politically relevant theme of Barack's tweets carries over into a more complete demonstration when analyzing these bigrams, with examples revealed such as "climate change", "health care", "gun violence", and "paris agreement". With Michelle, our theory of self-referencing is somewhat justified by her most common bigram being "michelle obama". Aditionally, an activism towards more auxiliary political issues is revealed through this list by entries such as "military families" and "girls education".

# Proportion of retweets

```{r}
#| echo: false
#| message: false
#| warning: false
tweets |>
  mutate(tweet_type = ifelse(str_detect(text, "^RT"), "retweet", "initial tweet")) |>
  group_by(person) |>
  summarise(prop_retweet = mean(tweet_type == "retweet")) |>
  ggplot(aes(x = person, y = prop_retweet)) +
  geom_col() +
  labs(
    title = "Proportion of retweets from\nBarack and Michelle Obama",
    subtitle = "(between May 2015 and November 2016)",
    caption = "**only contains words from the dplyr 'words' list**",
    x = "Person",
    y = "Proportion of retweets"
  )
```

Some of the different tweeting habits between Michelle and Barack could potentially be explained by how often either one of them send out a retweet versus sending out an original tweet. By detecting which of each person's tweets start with the letters "RT" we can identify if any given particular tweet was a retweet or not, with our resulting bar graph showing a significant difference in proportion of retweets between Barack and Michelle.

# Sentiments of tweets

```{r}
#| echo: false
#| message: false
#| warning: false
bing_sentiments <- get_sentiments(lexicon = "bing")

barack_words |>
  inner_join(bing_sentiments) |>
  count(year, month, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |>
  mutate(sentiment = (positive - negative) / (positive + negative)) |>
  ggplot(aes(x = month, y = sentiment, fill = year)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~year, ncol = 2, scales = "free_x") +
  labs(
    title = "Proportion of positive vs. negative sentiments\nin Barack Obama's tweets by month",
    subtitle = "(between May 2015 and November 2016)",
    caption = "**only contains words from the dplyr 'words' list**",
    x = "Month",
    y = "Proportion of positive vs. negative sentiment"
  )
```

For the remaining two charts, we will solely be focusing on Barack's tweeting history to see if we can garner any further insights about his final 20 months as president. First we will be looking at a measure of "net proportional positivity" for each of our recorded months by appending either a positive and negative sentiment to a select subset of our words. We then add together the net sum of each sentiment (assuming that a positive word has a +1 value and a negative word has a -1 value) and dividing that number by the total number of words utilized in the calculation. What we find is an oddly positive sentiment from Baracks tweets across every single month with no net proportional sentiment dipping under 0.5 and many of them sitting at a proportion of 1.0. This could support the driven and fearless character that President Obama seems to have built his campaign off of as shown by his word choice in our previous charts.

# Mood words of tweets

```{r}
#| echo: false
#| message: false
#| warning: false
nrc_sentiments <- get_sentiments("nrc")

barack_words |>
  inner_join(nrc_sentiments) |>
  count(year, month, sentiment) |>
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) |>
  mutate(
    sum_words = anger + anticipation + fear + joy + negative + positive + surprise + sadness + trust,
    anger = anger / sum_words,
    anticipation = anticipation / sum_words,
    fear = fear / sum_words,
    joy = joy / sum_words,
    positive = positive / sum_words,
    negative = negative / sum_words,
    sadness = sadness / sum_words,
    surprise = surprise / sum_words,
    trust = trust / sum_words
  ) |>
  pivot_longer(
    cols = c(anger:trust),
    names_to = "sentiment",
    values_to = "prop"
  ) |>
  ggplot() +
  geom_col(aes(x = month, y = prop, fill = sentiment), position = "stack") +
  facet_wrap(~year, ncol = 2, scales = "free_x") +
  labs(
    title = "Distribution of different mood words in Barack Obama's tweets by month",
    subtitle = "(between May 2015 and November 2016)",
    caption = "**only contains words from the dplyr 'words' list**",
    x = "Month",
    y = "Proportion"
  )
```

Finally, we analyze a plethora of different moods of words by appending a much broader list of emotions to our list of words. We derive the same process from our previous chart by taking the proportion of each emotion used in the words of Barack's tweets and dividing it by the total amount of words tweeted within that month. The results show a surprisingly unwavering proportion of each emotion for each month, regardless of how many total words were tweeted out during that month, which could possibly allude to an unwavering character not easily influenced by outside events or conflicts.